---
title: "Noise Pollution and Health"
author: "Lucas Schiffer"
date: "May 22, 2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggmap)
load("bcnp.rda")
load("nynp.rda")
load("sfnp.rda")
dv <- c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
```

## What is Noise Pollution?
"... an unwanted sound which gets dumped into [the] environment without regard... to the adverse effects it may be having. It adversely affects the health of a person and produces ill effects in living and non-living things." - Tripathy DP. Noise Pollution. New Delhi: A.P.H. Pub. Corp.; 1999.

## How Does It Impact Health?
- Stress Related Effects
- Teratogenic Effects
- Learning Ability
- Personal Safety
- Sleep Disruption

## Trends In 3 US Cities In 2015
- Baltimore, MD
- New York, NY
- San Fransisco, CA

- Summary
- Location
- Day of Week
- Season of Year

## Baltimore, MD Summary
```{r eval=FALSE}
gcll <- geocode("Baltimore, MD")
gcbb <- c(left = gcll$lon*1.2, bottom = gcll$lat*0.8, right = gcll$lon*0.8, top = gcll$lat*1.2)
gcgm <- get_stamenmap(gcbb, zoom = 7, maptype = "toner-background")
ggmap(gcgm, extent = "device") + geom_point(aes(x = lon, y = lat), data = gcll, colour = "red", size = 5)
```
- GDP (2014 Est.) <- $173,747 (PPP, $M)
- GDP / Capita (2014 Est.) <- $62,291 (PPP, $M)
- Population (2014 Est.) <- 2,789,289
- Noise Complaints (2015 Est.) <- 198

## Baltimore, MD Location
```{r eval=FALSE}
llld <- subset(bcnp, select = c("longitude", "latitude"))
colnames(llld) <- c("lon", "lat")
llld$lon <- as.numeric(llld$lon)
llld$lat <- as.numeric(llld$lat)
qmplot(lon, lat, data = llld, geom = "blank", zoom = 15, maptype = "toner-background") +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", alpha = .3, color = NA) +
  scale_fill_gradient2("", low = "white", mid = "yellow", high = "red", guide = FALSE)
```

## Baltimore, MD Day of Week
```{r}
dotw_ds <- bcnp$date
dotw_df <- as.data.frame(table(as.character(lapply(dotw_ds, function(x) weekdays(as.Date(as.character(x), "%m/%d/%Y"))))))
dotw_cv <- as.character(lapply(dotw_ds, function(x) weekdays(as.Date(as.character(x), "%m/%d/%Y"))))
qplot(dotw_cv, geom = "bar", main = "", xlab = "", ylab = "") + coord_flip() + theme_minimal() + scale_x_discrete(limits=rev(dv)) #+ scale_x_discrete(limits=dotw_df[order(dotw_df[,2]),]$Var1)
```

## Baltimore, MD Season of Year
```{r eval = FALSE}
moty_ds <- bcnp$date
moty_df <- as.data.frame(table(as.character(lapply(moty_ds, function(x) months(as.Date(as.character(x), "%m/%d/%Y"))))))
moty_cv <- as.character(lapply(moty_ds, function(x) months(as.Date(as.character(x), "%m/%d/%Y"))))
qplot(moty_cv, geom = "bar", main = "", xlab = "", ylab = "") + scale_x_discrete(limits=sort(factor(moty_df$Var1, levels = month.name))) + theme_minimal() + theme(axis.text.x=element_text(angle = -45, hjust = 0))
```

## New York, NY Summary
```{r eval=FALSE}
gcll <- geocode("New York, NY")
gcbb <- c(left = gcll$lon*1.2, bottom = gcll$lat*0.8, right = gcll$lon*0.8, top = gcll$lat*1.2)
gcgm <- get_stamenmap(gcbb, zoom = 7, maptype = "toner-background")
ggmap(gcgm, extent = "device") + geom_point(aes(x = lon, y = lat), data = gcll, colour = "red", size = 5)
```
- GDP (2014 Est.) <- $1,403,463 (PPP, $M)
- GDP / Capita (2014 Est.) <- $69,915 (PPP, $M)
- Population (2014 Est.) <- 20,073,930
- Noise Complaints (2015 Est.) <- 176,691

## New York, NY Location
```{r eval=FALSE}
llld <- subset(nynp, select = c("longitude", "latitude"))
colnames(llld) <- c("lon", "lat")
llld$lon <- as.numeric(llld$lon)
llld$lat <- as.numeric(llld$lat)
qmplot(lon, lat, data = llld, geom = "blank", zoom = 15, maptype = "toner-background") +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", alpha = .3, color = NA) +
  scale_fill_gradient2("", low = "white", mid = "yellow", high = "red", guide = FALSE)
```

## New York, NY Day of Week
```{r}
dotw_ds <- nynp$date
dotw_df <- as.data.frame(table(as.character(lapply(dotw_ds, function(x) weekdays(as.Date(as.character(x), "%m/%d/%Y"))))))
dotw_cv <- as.character(lapply(dotw_ds, function(x) weekdays(as.Date(as.character(x), "%m/%d/%Y"))))
qplot(dotw_cv, geom = "bar", main = "", xlab = "", ylab = "") + coord_flip() + theme_minimal() + scale_x_discrete(limits=rev(dv)) #+ scale_x_discrete(limits=dotw_df[order(dotw_df[,2]),]$Var1)
```

## New York, NY Season of Year
```{r eval = FALSE}
moty_ds <- nynp$date
moty_df <- as.data.frame(table(as.character(lapply(moty_ds, function(x) months(as.Date(as.character(x), "%m/%d/%Y"))))))
moty_cv <- as.character(lapply(moty_ds, function(x) months(as.Date(as.character(x), "%m/%d/%Y"))))
qplot(moty_cv, geom = "bar", main = "", xlab = "", ylab = "") + scale_x_discrete(limits=sort(factor(moty_df$Var1, levels = month.name))) + theme_minimal() + theme(axis.text.x=element_text(angle = -45, hjust = 0))
```


## San Fransisco, CA Summary
```{r  eval=FALSE}
gcll <- geocode("San Francisco, CA")
gcbb <- c(left = gcll$lon*1.2, bottom = gcll$lat*0.8, right = gcll$lon*0.8, top = gcll$lat*1.2)
gcgm <- get_stamenmap(gcbb, zoom = 7, maptype = "toner-background")
ggmap(gcgm, extent = "device") + geom_point(aes(x = lon, y = lat), data = gcll, colour = "red", size = 5)
```
- GDP (2014 Est.) <- $331,024 (PPP, $M)
- GDP / Capita (2014 Est.) <- $72,390 (PPP, $M)
- Population (2014 Est.) <- 4,572,807
- Noise Complaints (2015 Est.) <- 792

## San Fransisco, CA Location
```{r eval=FALSE}
llld <- subset(sfnp, select = c("longitude", "latitude"))
colnames(llld) <- c("lon", "lat")
llld$lon <- as.numeric(llld$lon)
llld$lat <- as.numeric(llld$lat)
qmplot(lon, lat, data = llld, geom = "blank", zoom = 15, maptype = "toner-background") +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", alpha = .3, color = NA) +
  scale_fill_gradient2("", low = "white", mid = "yellow", high = "red", guide = FALSE)
```

## San Fransisco, CA Day of Week
```{r}
dotw_ds <- sfnp$date
dotw_df <- as.data.frame(table(as.character(lapply(dotw_ds, function(x) weekdays(as.Date(as.character(x), "%m/%d/%Y"))))))
dotw_cv <- as.character(lapply(dotw_ds, function(x) weekdays(as.Date(as.character(x), "%m/%d/%Y"))))
qplot(dotw_cv, geom = "bar", main = "", xlab = "", ylab = "") + coord_flip() + theme_minimal() + scale_x_discrete(limits=rev(dv)) #+ scale_x_discrete(limits=dotw_df[order(dotw_df[,2]),]$Var1)
```

## San Fransisco, CA Season of Year
```{r eval = FALSE}
moty_ds <- c(sfnp$date, "01/01/2015", "02/01/2015", "03/01/2015", "04/01/2015", "05/01/2015", "06/01/2015", "07/01/2015")
moty_df <- as.data.frame(table(as.character(lapply(moty_ds, function(x) months(as.Date(as.character(x), "%m/%d/%Y"))))))
moty_cv <- as.character(lapply(moty_ds, function(x) months(as.Date(as.character(x), "%m/%d/%Y"))))
qplot(moty_cv, geom = "bar", main = "", xlab = "", ylab = "") + scale_x_discrete(limits=sort(factor(moty_df$Var1, levels = month.name))) + theme_minimal() + theme(axis.text.x=element_text(angle = -45, hjust = 0))
```

## Compartive Noise Polution Per Capita
- Baltimore, MD <- 7.10 / 100,000
- New York, NY <- 880.20 / 100,000
- San Fransisco, CA <-  17.32 / 100,000

## References
```{r eval = FALSE}
citation("base")
citation("downloader")
citation("data.table")
citation("stringr")
citation("ggmap")
```
